```{r}
library(tidyverse)

theme_set(theme_classic(base_size = 20)) 
```

using lapply to make a list of files (save as object)
```{r}
files=lapply(list.files(path = "movies",full.names = F, pattern = ".txt"),function(file_name){
  read_delim(paste0("movies/",file_name),delim = ",")
})
names(files)=str_replace(string=list.files(path = "movies",full.names = F, pattern = ".txt"),pattern = ".txt",replacement = "")
names(files)
movies_joined=full_join(files[["movies_imdb"]],files[["movies_rottentom"]],by=c("movie_title"="title"))


## subset of the data
set.seed(100) ##this allows us to work with the same random sample
movies_subset=movies_joined %>%
  select(budget,title_year,movie_title,duration,imdb_score,metacritic,rotten_tomatoes) %>%
  na.omit() %>%
  sample_n(100)
```


plotting different fits
```{r}
ggplot(movies_subset,aes(x=imdb_score,y=rotten_tomatoes))+
  geom_point()+
  geom_smooth(method="lm") +
  ggtitle("linear model")

ggplot(movies_subset,aes(x=imdb_score,y=rotten_tomatoes))+
  geom_point()+
  geom_smooth(formula = "y~exp(x)") +
  ggtitle("exponential: y~exp(x)")

ggplot(movies_subset,aes(x=imdb_score,y=rotten_tomatoes))+
  geom_point()+
  geom_smooth(formula = "y~poly(x,2)") +
  ggtitle("quadratic: y~poly(x,2)")
```


#Mediation using example mediation data
```{r}
mediation_data=read_csv("mediation_data.csv")

## plot all 3
library(gridExtra)
grid.arrange(ggplot(mediation_data,aes(x=X,y=Y))+geom_point(),
             ggplot(mediation_data,aes(x=X,y=M))+geom_point(),
             ggplot(mediation_data,aes(x=M,y=Y))+geom_point(),
             nrow=1)
```

model them to show the 3 pre-conditions are met: 
```{r} 
library(broom)
model.0=lm(Y~X,mediation_data)
tidy(model.0)
model.M=lm(M~X,mediation_data)
tidy(model.M)
model.Y=lm(Y~X+M,mediation_data)
tidy(model.Y)


## proportion mediation: IDE/c', which is also (c-c')/c
(coef(model.0)[2]-coef(model.Y)[2])/coef(model.0)[2]
```

bootstrap test for mediation
```{r} 
nboot=1000
boot_list=lapply(1:nboot,function(iteration){
  boot_data=sample(mediation_data,replace = T) ## resample with replacement
  m1=lm(Y~X,boot_data) ## model for c
  m2=lm(Y~X+M,boot_data)  ## model for c'
  c1=coef(m1)[2]
  cprime=coef(m2)[2]
  return(c(c1,cprime))})

boot_results=Reduce(rbind,boot_list) %>%
  as_tibble()
colnames(boot_results)=c("c","cprime")

boot_results

```

calculate indirect effects and plot
```{r}
boot_results=boot_results %>% 
  mutate(ide=c - cprime)

boot_results %>% 
  gather(key=path,value=effect) %>%
  ggplot(aes(x=effect,fill=path))+
  geom_histogram(position = "dodge")
```


estimate the signficance of the mediation 
(note that this is a strong mediation effect)
```{r}
boot_results %>% 
  summarize(sum(ide<0)/n())

## complete mediation? 
boot_results %>% 
  summarize(median(cprime))
```


using the `mediation` package
```{r}
library(mediation)

model.M=lm(M~X,mediation_data)
tidy(model.M)
model.Y=lm(Y~X+M,mediation_data)
tidy(model.Y)

med_model=mediate(model.M, model.Y, treat='X', mediator='M',
                   boot=TRUE, sims=1000)

summary(med_model)
```

#Visualizing multiple regression.
First we explore the data and plot a few associations
```{r}
smoking=read_csv("smoking_data.csv")

library(skimr)
smoking %>% skim()

library(ggbeeswarm)
smoking %>%
  ggplot(aes(x=smoke,y=lung_capacity))+
  geom_beeswarm()
```

Being exposed to smoking increases lung capacity?
```{r}
smoking %>%
  ggplot(aes(x=age,y=lung_capacity))+
  geom_point()

smoking %>%
  ggplot(aes(x=age,y=lung_capacity,col=smoke))+
  geom_beeswarm()

smoking %>%
  ggplot(aes(x=age,y=lung_capacity,col=smoke))+
  geom_beeswarm()+
  stat_smooth(method="lm")
```

smoking has an negative effect when we control for age. 
```{r}
model1=lm(formula = lung_capacity ~ smoke, data = smoking)

model2=lm(formula = lung_capacity ~ age + smoke, data = smoking)

tidy(model1)
tidy(model2)
```

two ways to visualize this (what's another way?)
```{r}
#1 plot smoking variable vs. residuals of a model excluding smoking
## note that this doesn't always work well for visualization, but at least the x-axis is interpretable
y_given_x1=resid(lm(lung_capacity ~ age, data = smoking))
tibble(smoke=smoking$smoke,y_given_x1=y_given_x1) %>%
  ggplot(aes(x=smoke,y=y_given_x1))+
  geom_beeswarm(alpha=0.1)

#2 partial residuals, this usually works, but the x-axis can be less interpretable. 
smoking$smoke_integer=as.numeric(as.factor(smoking$smoke))
head(smoking$smoke_integer)
x2_given_x1=resid(lm(smoke_integer ~ age, data = smoking))
y_given_x1=resid(lm(lung_capacity ~ age, data = smoking))

tibble(x2_given_x1,y_given_x1) %>%
  ggplot(aes(x=x2_given_x1,y=y_given_x1))+
  geom_point()+
  stat_smooth(method="lm")

```

