```{r}
library(tidyverse)

theme_set(theme_classic(base_size = 20)) 
```

using lapply to make a list of files (save as object)
```{r}
files=lapply(list.files(path = "movies",full.names = F, pattern = ".txt"),function(file_name){
  read_delim(paste0("movies/",file_name),delim = ",")
})
names(files)=str_replace(string=list.files(path = "movies",full.names = F, pattern = ".txt"),pattern = ".txt",replacement = "")
names(files)
movies_joined=full_join(files[["movies_imdb"]],files[["movies_rottentom"]],by=c("movie_title"="title"))

```

bootstrapping for a confidence interval on a mean
```{r}
movies_action=files[["movies_imdb"]] %>% 
  mutate(action=str_detect(genres,"Action")) %>%
  select(action,title_year,imdb_score)

set.seed(10) ##this allows us to work with the same random sample
movies_action_sample <- movies_action %>% 
  filter(!is.na(imdb_score)) %>%
  sample_n(50)

movies_action_sample %>% 
    specify(response=imdb_score) %>% 
    generate(reps = 100,type="bootstrap") %>%
    calculate(stat="mean")%>%
    visualize()

movies_action_sample %>% 
    specify(response=imdb_score) %>% 
    generate(reps = 1000,type="bootstrap") %>%
    calculate(stat="mean") %>%
    get_ci(level = 0.95, type = "percentile")
```

compare permutations to bootstrapping
```{r}
movies_action=files[["movies_imdb"]] %>% 
  mutate(action=str_detect(genres,"Action")) %>%
  select(action,title_year,imdb_score)

set.seed(10) ##this allows us to work with the same random sample
movies_action_sample <- movies_action %>% 
  group_by(action) %>%
  sample_n(25) %>% 
  ungroup()

permuted=movies_action_sample %>% 
  specify(formula = imdb_score ~ action) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 100,type="permute") %>% 
  calculate(stat = "diff in means", order = c("TRUE", "FALSE"))
permuted$random="permutation"

bootstrapped=movies_action_sample %>% 
  specify(formula = imdb_score ~ action) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 100,type="bootstrap") %>% 
  calculate(stat = "diff in means", order = c("TRUE", "FALSE"))
bootstrapped$random="bootstrap"

combined=rbind(bootstrapped,permuted) #combine them

ggplot(combined,aes(x=stat))+
  geom_histogram(fill="grey25")+
  facet_wrap(~random,nrow = 2)
```


bootstrapping regression coefficients 
first we can estimate a regression coefficient
```{r}
set.seed(100) ##this allows us to work with the same random sample
movies_subset=movies_joined %>%
  select(budget,title_year,movie_title,duration,imdb_score,metacritic,rotten_tomatoes) %>%
  na.omit() %>%
  sample_n(100)

lm(rotten_tomatoes~imdb_score,movies_subset)

library(broom)  ## to view tidy summaries
tidy(lm(rotten_tomatoes~imdb_score,movies_subset))

```

bootstrap or permute regression coefficients
```{r}
movies_subset %>%
    specify(rotten_tomatoes~imdb_score) %>%
    generate(reps = 1000, type = "bootstrap") %>% 
    calculate(stat = "slope") %>%
    visualize()

movies_subset %>%
    specify(rotten_tomatoes~imdb_score) %>%
    hypothesize(null = "independence") %>%
    generate(reps = 1000, type = "permute") %>%
    calculate(stat = "slope") %>%
    visualize()
```


but this can't be expanded to multiple predictors using `infer`
So we use the `mosiac` package
```{r}
library(mosaic)
lm_boot=do(100) * lm(rotten_tomatoes~imdb_score, 
                     data = sample(movies_subset,replace=T))
confint(lm_boot)
```

what happens if we add in multiple variables?
```{r}
lm_boot=do(1000) * lm(rotten_tomatoes~imdb_score + budget, 
                     data = sample(movies_subset,replace=T))

lm_boot %>% head()

confint(lm_boot) 

ggplot(movies_subset,aes(x=imdb_score,y=rotten_tomatoes))+
  geom_abline(intercept = lm_boot$Intercept,
              slope=lm_boot$imdb_score,col="grey50",lwd=3)+
  geom_point()+
  geom_smooth(method="lm",se=F)
```

plot all of the bootstrap iterations, what's another way to plot this?
```{r}
ggplot(movies_subset,aes(x=imdb_score,y=rotten_tomatoes))+
  geom_abline(intercept = lm_boot$Intercept,
              slope=lm_boot$imdb_score,col="grey70",lwd=3)+
  geom_point()+
  geom_abline(intercept=-54.8,slope=15.1,col="blue",lwd=2)
```