start your engines (load your packages)
```{r}
library(tidyverse)

theme_set(theme_classic(base_size = 20)) 
```

For loop to read in files (first we list the files)
```{r}
list.files(path = "movies",full.names = F)

list.files(path = "movies",full.names = F, pattern = ".txt")

files=list.files(path = "movies",full.names = F, pattern = ".txt")
for (file_name in files){
  print(file_name)}

for (file_name in files){
  print(paste0("movies/",file_name))}
```

Now that we know what it does, we can read them in. whoops
```{r}
for (file_name in files){
 file=read_delim(paste0("movies/",file_name),delim = ",")
 }
```


can we assign the name of the files? using assign()
```{r}
for (file_name in files){
 assign(file_name,read_delim(paste0("movies/",file_name),delim = ","))
 }
```

can we remove the ".txt" extension from the object name?
```{r}
for (file_name in list.files(path = "movies",full.names = F, pattern = ".txt")){
  file_n=str_replace(file_name,".txt","")
 assign(file_n,read_delim(paste0("movies/",file_name),delim = ","))
}

## clean up the extra objects
rm(file_n)
rm(file_name)
```


using lapply to make a list of files
```{r}
lapply(files,function(file_name){
  read_delim(paste0("movies/",file_name),delim = ",")
})
```

using lapply to make a list of files (save as object)
```{r}
files=lapply(list.files(path = "movies",full.names = F, pattern = ".txt"),function(file_name){
  read_delim(paste0("movies/",file_name),delim = ",")
})

View(files[[1]])
View(files[[2]])

names(files)
names(files)=list.files(path = "movies",full.names = F, pattern = ".txt")
names(files)
names(files)=str_replace(string=list.files(path = "movies",full.names = F, pattern = ".txt"),pattern = ".txt",replacement = "")
names(files)

View(files[["movies_imdb"]])
View(files[["movies_rottentom"]])
```

Sets
```{r}
intersect(movies_imdb$movie_title, movies_rottentom$title) %>% View()
union(movies_imdb$movie_title, movies_rottentom$title) %>% View()
setdiff(movies_imdb$movie_title, movies_rottentom$title) %>% View()
setdiff(movies_rottentom$title, movies_imdb$movie_title) %>% View()

movies_rottentom$title %in% movies_imdb$movie_title %>% View()
```

Let's look at some distributions. 
```{r}
movies_imdb=movies[["movies_imdb.txt"]]
ggplot(data=movies_imdb,aes(x=imdb_score))+
  geom_histogram()
```

what is a "bin"?
```{r}
ggplot(data=movies_imdb,aes(x=imdb_score))+
  geom_histogram(bins=10)
```

what about "bin width"?
```{r}
ggplot(data=movies_imdb,aes(x=imdb_score))+
  geom_histogram(binwidth = 0.1)
```

What about "bin width"?
```{r}
ggplot(data=movies_imdb,aes(x=imdb_score))+
  geom_histogram(binwidth = 0.1)
```

What about comparing multiple distributions?
```{r}
ggplot(data=movies_imdb,aes(x=imdb_score,fill=color))+
  geom_histogram(binwidth = 0.1)
```

stacked bar charts are confusing. what else can we do?
```{r}
ggplot(data=movies_imdb,aes(x=imdb_score,fill=color))+
  geom_histogram(binwidth = 0.1, position = "dodge")+
  scale_fill_brewer(palette = "Set1")

ggplot(data=movies_imdb,aes(x=imdb_score,col=color))+
  geom_freqpoly(binwidth = 0.1)+scale_color_brewer(palette = "Set1")
```

we have very different sample sizes here, which makes it hard to compare the two distributions. But we could use density plots. 
```{r}
movies_imdb %>% count(color)

ggplot(data=movies_imdb,aes(x=imdb_score,fill=color)) +
  geom_density()+scale_fill_brewer(palette = "Set1")
```

let's get rid of the NAs and make the density plots a bit transparent
```{r}
color_bw_movies=movies_imdb %>% filter(!is.na(color))

ggplot(data=color_bw_movies,aes(x=imdb_score,fill=color))+geom_density(alpha=0.5)+scale_fill_brewer(palette = "Set1")
```

Ohhhh - looks like B/W movies are higher rated (on average) compared to color. We can test this a couple of ways. 
```{r}
##T-test (can be paired or not)
t.test(color_bw_movies$imdb_score~color_bw_movies$color)

## Wilcoxon T-test (also known as a Mann-Whitney, can be paired or not)
wilcox.test(color_bw_movies$imdb_score~color_bw_movies$color)
```

Let's join the movie data and do some more data exploration
```{r}
movies_joined=full_join(movies_imdb,movies_rottentom,by=c("movie_title"="title"))
```

Let's make a  IMDB, rotten tomatoes, and metacritic scores from the same distribution?
```{r}
movies_joined %>% 
    select(movie_title,metacritic,imdb_score,rotten_tomatoes) %>%
    gather(key=source,value = score,c(metacritic,imdb_score,rotten_tomatoes)) %>%
    ggplot(aes(x=score,fill=source))+
    geom_density(alpha=0.5) +
    scale_fill_brewer(palette = "Set1")
```

THEY AREN'T ON THE SAME SCALE, LET'S FIX THAT
use the scale() function
```{r}
movies_joined %>% 
    select(movie_title,metacritic,imdb_score,rotten_tomatoes) %>%
    gather(key=source,value = score,c(metacritic,imdb_score,rotten_tomatoes)) %>%
    group_by(source) %>%
    mutate(score=scale(score)) %>%
    ggplot(aes(x=score,fill=source))+
    geom_density(alpha=0.5) +
    scale_fill_brewer(palette = "Set1")

```
