```{r}
library(tidyverse)

theme_set(theme_classic(base_size = 20)) 
```

using lapply to make a list of files (save as object)
```{r}
files=lapply(list.files(path = "movies",full.names = F, pattern = ".txt"),function(file_name){
  read_delim(paste0("movies/",file_name),delim = ",")
})
names(files)=str_replace(string=list.files(path = "movies",full.names = F, pattern = ".txt"),pattern = ".txt",replacement = "")
names(files)
movies_joined=full_join(files[["movies_imdb"]],files[["movies_rottentom"]],by=c("movie_title"="title"))


## subset of the data
set.seed(100) ##this allows us to work with the same random sample
movies_subset=movies_joined %>%
  select(budget,title_year,movie_title,duration,imdb_score,metacritic,rotten_tomatoes) %>%
  na.omit() %>%
  sample_n(100)
```


plotting different fits
```{r}
ggplot(movies_subset,aes(x=imdb_score,y=rotten_tomatoes))+
  geom_point()+
  geom_smooth(method="lm") +
  ggtitle("linear model")

ggplot(movies_subset,aes(x=imdb_score,y=rotten_tomatoes))+
  geom_point()+
  geom_smooth(formula = "y~exp(x)",method="lm") +
  ggtitle("exponential: y~exp(x)")

ggplot(movies_subset,aes(x=imdb_score,y=rotten_tomatoes))+
  geom_point()+
  geom_smooth(formula = "y~poly(x,2)",method="lm") +
  ggtitle("quadratic: y~poly(x,2)")
```

#Interactions

generate some fake data with a strong interaction
```{r}
set.seed(100)
a=mvrnorm(100,mu=c(5,5),Sigma = matrix(c(1,0.75,0.75,1),ncol=2))
b=mvrnorm(100,mu=c(5,5),Sigma = matrix(c(1,-0.75,-0.75,1),ncol=2))
a=as_tibble(a)
a$group="A"
b=as_tibble(b)
b$group="B"
df=rbind(a,b)

ggplot(df,aes(x=V1,y=V2))+
  geom_point() +
  stat_smooth(method="lm")

tidy(lm(V2~V1,data=df))
```

let's look at this in a better way
```{r}
ggplot(df,aes(x=V1,y=V2,col=group))+
  geom_point() +
  stat_smooth(method="lm")

tidy(lm(V2~V1*group,data=df))

```


#Mediation using example mediation data
```{r}
mediation_data=read_csv("mediation_data.csv")

## plot all 3
library(gridExtra)
grid.arrange(ggplot(mediation_data,aes(x=X,y=Y))+geom_point(),
             ggplot(mediation_data,aes(x=X,y=M))+geom_point(),
             ggplot(mediation_data,aes(x=M,y=Y))+geom_point(),
             nrow=1)
```

model them to show the 3 pre-conditions are met: 
```{r} 
library(broom)
model.0=lm(Y~X,mediation_data)
tidy(model.0)
model.M=lm(M~X,mediation_data)
tidy(model.M)
model.Y=lm(Y~X+M,mediation_data)
tidy(model.Y)


## proportion mediation: IDE/c', which is also (c-c')/c
(coef(model.0)[2]-coef(model.Y)[2])/coef(model.0)[2]
```

bootstrap test for mediation
```{r} 
nboot=1000
boot_list=lapply(1:nboot,function(iteration){
  boot_data=sample_n(mediation_data,replace = T,size=nrow(mediation_data)) ## resample with replacement
  m1=lm(Y~X,boot_data) ## model for c
  m2=lm(Y~X+M,boot_data)  ## model for c'
  c1=coef(m1)[2]
  cprime=coef(m2)[2]
  return(c(c1,cprime))
  }
  )

boot_results=Reduce(rbind,boot_list) %>%
  as_tibble()
colnames(boot_results)=c("c","cprime")

boot_results

```

calculate indirect effects and plot
```{r}
boot_results=boot_results %>% 
  mutate(ide=c - cprime)

boot_results %>% 
  gather(key=path,value=effect) %>%
  ggplot(aes(x=effect,fill=path))+
  geom_histogram(position = "dodge")
```


estimate the signficance of the mediation 
(note that this is a strong mediation effect)
```{r}
boot_results %>% 
  summarize(sum(ide<0)/n())

## complete mediation? 
boot_results %>% 
  summarize(median(cprime))
```


using the `mediation` package
```{r}
library(mediation)

model.M=lm(M~X,mediation_data)
tidy(model.M)
model.Y=lm(Y~X+M,mediation_data)
tidy(model.Y)

med_model=mediate(model.M, model.Y, treat='X', mediator='M',
                   boot=TRUE, sims=1000)

summary(med_model)
```

