---
title: "Lecture 10 in class activity"
date: "5/2/2019"
output: html_document
---

#  Multiple regression and interactions

load in the data
```{r}
library(car)
library(tidyverse)
theme_set(theme_classic(base_size = 20)) 

data("Salaries")
head(Salaries)
```

Using the Salaries dataset from the car package:

1. filter these data to observations less than 25 years since the PhD. 
2. plot salary as a function of years since PhD, discipline, and sex (in one or two plots)
```{r}

```


Fit a linear model predicting salary as a function of sex + yrs.since.phd + discipline.  

Look at the coefficients using tidy(). 

```{r}

```

In this model, how much is salary expected to increase for each year after the phd?

Does the effect of years since phd depend on the sex of the professor? 

Test this by fitting a model with an interaction between sex and yrs.since.phd
```{r}

```

Using the diamonds data from the  ggplot2 package, filter this dataset to include only diamonds with Fair and Ideal cuts, and store the result in a new tibble called d.small.  The cut variable is stored as an ordered factor.  Change this to an unordered factor by first converting it to character, and then back to factor: as.character(), as.factor()
```{r}
data("diamonds")
diamonds = diamonds %>% filter(cut %in% c("Ideal","Fair"))
diamonds$cut = as.character(diamonds$cut)
diamonds$cut = as.factor(diamonds$cut)
```


plot mapping price to y and carat to x.  
Add points and a geom_smooth(method = 'lm') and facet the plot by the cut variable
```{r}

```

Fit a model predicting price as a function of carat, cut and the interaction of those two variables.  Is the interaction significant?
```{r}


```


Look at the levels of the cut variable using the levels() function.
What is the reference level?

```{r}


```


What is the expected increase in price for a one unit increase in carat at the reference level of cut?

Now relevel the cut variable so that Ideal is the reference level.
```{r}
diamonds$cut = relevel(diamonds$cut, ref = 'Ideal')
```
 
Fit the model again and assess the new estimate for change in price with a one carat increase.
```{r}
d.model2 = lm(price~carat*cut, data = diamonds)
tidy(d.model2)
```

Make the following data frame with two cuts and predict their prices using d.model2 and the predict() function
```{r}
df = data.frame(cut = c('Ideal','Fair'), carat = c(3,3))

df$price = predict(d.model2, df)
```

Add points to the ggplot you made previously with these predicted values shown in red on the plot.  
Map the carat variable to X, and the predicted price to Y
```{r}
ggplot(diamonds, aes(x = carat, y = price))+
     geom_point(alpha=0.1) + 
     geom_smooth(method = 'lm') + 
     facet_grid(~cut)+ 
    geom_point(data = df, aes(x = carat, y = price), color = "red")
```

Do the points fall where you expect them to?
