
```{r}
library(tidyverse)
```

read in the files
```{r}
movies_imdb=read_delim("movies/movies_imdb.txt",delim=",")
movies_rottentom=read_delim("movies/movies_rottentom.txt",delim=",")
```

## Including Plots

You can also embed plots, for example:

```{r}
movies_imdb %>% select(movie_title,title_year,duration,imdb_score)
```

```{r}
imdb_minimal=movies_imdb %>% select(movie_title,title_year,duration,imdb_score)
```


```{r}
imdb_minimal=movies_imdb %>% select(movie_title,title_year,duration,imdb_score)

joined=full_join(imdb_minimal,movies_rottentom,by=c("movie_title"="title"))
```


ggplot data
```{r}
plot_imdb = ggplot(movies_imdb)
summary(plot_imdb)
```

ggplot aesthetics
```{r}
plot_imdb = ggplot(movies_imdb) + aes(x=title_year,y=imdb_score)
summary(plot_imdb)
```

Add layers to a ggplot object with +
```{r}
plot_imdb = ggplot(movies_imdb)
plot_imdb = plot_imdb + aes(x=title_year,y=imdb_score)
summary(plot_imdb)
```

ggplot geoms
```{r}
plot_imdb = plot_imdb + geom_point()
summary(plot_imdb)
plot_imdb
```

back to the slides for a second.

some nice default themes 
```{r}
plot_imdb = plot_imdb + theme_classic(base_size = 20) 
plot_imdb
```

set it globally
```{r}
theme_set(theme_classic(base_size = 20)) 
```

and axis labels
```{r}
plot_imdb + xlab("movie release year") + ylab("IMDB score")
plot_imdb + xlab("movie release\n(year)") + ylab("IMDB score")
plot_imdb=plot_imdb + xlab("movie release year") + ylab("IMDB score")
```

ggplot scale
```{r}
plot_imdb + scale_y_continuous(limits=c(0,10))+scale_x_continuous(limits=c(1975,2010))
```

ggplot statistics
Note: every stat has a default geom, every geom has a default stat
```{r}
plot_imdb + stat_smooth()
plot_imdb + stat_smooth(method="lm")
plot_imdb + stat_smooth(method="lm",se=F)
```

Overplotting is a problem, how can we summarize?
```{r}
plot_imdb = ggplot(movies_imdb) + aes(x=title_year,y=imdb_score)+geom_point(alpha=0.1)+ xlab("movie release year") + ylab("IMDB score")+theme_classic(base_size = 20) + ggtitle("alpha == 0.1")
plot_imdb
plot_imdb = ggplot(movies_imdb) + aes(x=title_year,y=imdb_score)+geom_point(alpha=0.01)+ xlab("movie release year") + ylab("IMDB score")+theme_classic(base_size = 20) + ggtitle("alpha == 0.01")
plot_imdb
```

another way
```{r}
ggplot(movies_imdb) + aes(x=title_year,y=imdb_score)+xlab("movie release year") + ylab("IMDB score")+theme_classic(base_size = 20) + geom_bin2d()
```

another way
```{r}
ggplot(movies_imdb) + aes(x=title_year,y=imdb_score)+xlab("movie release year") + ylab("IMDB score")+theme_classic(base_size = 20) + geom_count(alpha=0.1)
```

another way
```{r}
ggplot(movies_imdb) + aes(x=title_year,y=imdb_score)+xlab("movie release year") + ylab("IMDB score")+geom_point()+theme_classic(base_size = 20) + geom_quantile()
```

Let's change the color of the points
Aesthetics: mapping vs. setting
```{r}
imdb_plot=ggplot(movies_imdb) + aes(x=title_year,y=imdb_score)+xlab("movie release year") + ylab("IMDB score")+theme_classic(base_size = 20)

imdb_plot+ geom_point(col="green")

imdb_plot+ geom_point(aes(col="green"))
```

So how would we color them by their total budget?
```{r}
imdb_plot=ggplot(movies_imdb) + aes(x=title_year,y=imdb_score)+xlab("movie release year") + ylab("IMDB score")+theme_classic(base_size = 20)

imdb_plot+geom_point(aes(col=budget))
```

Back to stats, how about plots for color and b/w movies?
```{r}
imdb_plot=ggplot(movies_imdb) + aes(x=title_year,y=imdb_score)+xlab("movie release year") + ylab("IMDB score")+theme_classic(base_size = 20)

imdb_plot+geom_point(aes(col=budget))
```

piping the tidyverse right into ggplots
```{r}
movies_imdb %>% group_by(title_year) %>% summarize(min_score=min(imdb_score),max_score=max(imdb_score),mean_score=mean(imdb_score)) %>% ggplot(aes(x=title_year,y=mean_score,ymin=min_score,ymax=max_score))+geom_errorbar()+geom_point()
```
