```{r}
library(tidyverse)

theme_set(theme_classic(base_size = 20)) 
```


using lapply to make a list of files (save as object)
```{r}
files=lapply(list.files(path = "movies",full.names = F, pattern = ".txt"),function(file_name){
  read_delim(paste0("movies/",file_name),delim = ",")
})

files[[1]]
files[[2]]

names(files)
names(files)=list.files(path = "movies",full.names = F, pattern = ".txt")
names(files)
names(files)=str_replace(string=list.files(path = "movies",full.names = F, pattern = ".txt"),pattern = ".txt",replacement = "")
names(files)

files[["movies_imdb"]]
files[["movies_rottentom"]]
```

Let's join the movie data and do some more data exploration
```{r}
movies_joined=full_join(files[["movies_imdb"]],files[["movies_rottentom"]],by=c("movie_title"="title"))
```


Another way to compare distributions: Quantile-Quantile (Q-Q) plots
We can compare against a normal distribution
```{r}
qqnorm(movies_joined$rotten_tomatoes)
## after z-scoring (to get on same scale) 
qqnorm(scale(movies_joined$rotten_tomatoes),main = "Rotten Tomatoes")
## adding a line of x=y: slope (b) = 1 and intercept (a) = 0
abline(a=0,b=1,col="red",lwd=3)
```

Let's do a quick quiz: which dataset is more normally distributed? 
IMDB, rotten tomato, or metacritic scores? 

QQ plots of two distributions
```{r}
qqplot(movies_joined$imdb_score,movies_joined$rotten_tomatoes)

qqplot(movies_joined$imdb_score,movies_joined$metacritic)
```

Q-Q plots from the same distribution fall on the x=y line (slope=1, intercept=0) 
```{r}
b=rnorm(1000) ## randomly draw 1000 samples from a normal distribution
a=rnorm(1000) 
qqplot(a,b)
abline(0,1,col="red",lwd=5)
```

Challenge: let's plot the eCDFs for 1000 draws of a normal distribution, log-normal, and uniform. 

Do action movies have a higher IMDB score? 
```{r}
movies_imdb = files[["movies_imdb"]] %>% 
  mutate(action=str_detect(genres,"Action"))

ggplot(movies_imdb,aes(x=imdb_score,fill=action))+
  geom_density(alpha=0.5)+
  scale_fill_brewer(palette = "Set1")
``` 

Correlations among variables
```{r}
movies_joined %>% 
  select(imdb_score,rotten_tomatoes,metacritic,budget) %>%
  na.omit() %>%
  cor()
``` 


Using corrgram (or corrplot, whichever you prefer)
```{r}
library(corrplot) # or corrgram, up to you
movies_joined %>% 
    select(imdb_score,rotten_tomatoes,metacritic,budget) %>%
    na.omit() %>%
    cor() %>% 
    corrplot(method = "ellipse")
```


Spotting outliers
```{r}
files[["movies_imdb"]] %>% 
  ggplot(aes(x=budget))+
  geom_histogram(bins=100)

```


bringing them in closer (log scale)
```{r}
files[["movies_imdb"]]  %>% 
  ggplot(aes(x=budget))+
  geom_histogram(bins=100)+
  scale_x_log10()
```


plotting multiple distributions
```{r}
library(ggridges)

movies_imdb %>% 
  mutate(action=str_detect(genres,"Action")) %>%
  ggplot(aes(x=budget,y=action))+
  geom_density_ridges(colour="white",fill="steelblue4")+
  theme_classic(base_size=10) +
  scale_x_log10()
```


it can scale to many distributions
```{r}

movies_imdb %>% 
  ggplot(aes(x=factor(title_year),y=budget))+
  geom_violin(fill="steelblue4")+
  theme_classic(base_size=10) +
  scale_y_log10()

library(ggridges)

movies_imdb %>% 
  ggplot(aes(x=budget,y=factor(title_year)))+
  geom_density_ridges(colour="white",fill="steelblue4")+
  theme_classic(base_size=10) +
  scale_x_log10()

```
