start your engines (load your packages)
```{r}
library(tidyverse)

theme_set(theme_classic(base_size = 20)) 
```

For loop to read in files (first we list the files)
```{r}
list.files(path = "movies",full.names = F)

list.files(path = "movies",full.names = F, pattern = ".txt")

files=list.files(path = "movies",full.names = F, pattern = ".txt")
for (file_name in files){
  print(file_name)}

for (file_name in files){
  print(paste0("movies/",file_name))}
```

Now that we know what it does, we can read them in. whoops
```{r}
for (file_name in files){
 file=read_delim(paste0("movies/",file_name),delim = ",")
 }
```


can we assign the name of the files? using assign()
```{r}
for (file_name in files){
 assign(file_name,read_delim(paste0("movies/",file_name),delim = ","))
 }
```

can we remove the ".txt" extension from the object name?
```{r}
for (file_name in list.files(path = "movies",full.names = F, pattern = ".txt")){
  file_n=str_replace(file_name,".txt","")
 assign(file_n,read_delim(paste0("movies/",file_name),delim = ","))
}

## clean up the extra objects
rm(file_n)
rm(file_name)
```


using lapply to make a list of files
```{r}
lapply(files,function(file_name){
  read_delim(paste0("movies/",file_name),delim = ",")
})
```

using lapply to make a list of files (save as object)
```{r}
files=lapply(list.files(path = "movies",full.names = F, pattern = ".txt"),function(file_name){
  read_delim(paste0("movies/",file_name),delim = ",")
})

View(files[[1]])
View(files[[2]])

names(files)
names(files)=list.files(path = "movies",full.names = F, pattern = ".txt")
names(files)
names(files)=str_replace(string=list.files(path = "movies",full.names = F, pattern = ".txt"),pattern = ".txt",replacement = "")
names(files)

View(files[["movies_imdb"]])
View(files[["movies_rottentom"]])
```

Sets
```{r}
intersect(movies_imdb$movie_title, movies_rottentom$title) %>% View()
union(movies_imdb$movie_title, movies_rottentom$title) %>% View()
setdiff(movies_imdb$movie_title, movies_rottentom$title) %>% View()
setdiff(movies_rottentom$title, movies_imdb$movie_title) %>% View()

movies_rottentom$title %in% movies_imdb$movie_title %>% View()
```

Let's look at some distributions. 
```{r}
movies_imdb=movies[["movies_imdb.txt"]]
ggplot(data=movies_imdb,aes(x=imdb_score))+
  geom_histogram()
```

what is a "bin"?
```{r}
ggplot(data=movies_imdb,aes(x=imdb_score))+
  geom_histogram(bins=10)
```

what about "bin width"?
```{r}
ggplot(data=movies_imdb,aes(x=imdb_score))+
  geom_histogram(binwidth = 0.1)
```

What about "bin width"?
```{r}
ggplot(data=movies_imdb,aes(x=imdb_score))+
  geom_histogram(binwidth = 0.1)
```

What about comparing multiple distributions?
```{r}
ggplot(data=movies_imdb,aes(x=imdb_score,fill=color))+
  geom_histogram(binwidth = 0.1)
```

stacked bar charts are confusing. what else can we do?
```{r}
ggplot(data=movies_imdb,aes(x=imdb_score,fill=color))+
  geom_histogram(binwidth = 0.1, position = "dodge")+
  scale_fill_brewer(palette = "Set1")

ggplot(data=movies_imdb,aes(x=imdb_score,col=color))+
  geom_freqpoly(binwidth = 0.1)+scale_color_brewer(palette = "Set1")
```

we have very different sample sizes here, which makes it hard to compare the two distributions. But we could use density plots. 
```{r}
movies_imdb %>% count(color)

ggplot(data=movies_imdb,aes(x=imdb_score,fill=color)) +
  geom_density()+scale_fill_brewer(palette = "Set1")
```

let's get rid of the NAs and make the density plots a bit transparent
```{r}
color_bw_movies=movies_imdb %>% filter(!is.na(color))

ggplot(data=color_bw_movies,aes(x=imdb_score,fill=color))+geom_density(alpha=0.5)+scale_fill_brewer(palette = "Set1")
```

Ohhhh - looks like B/W movies are higher rated (on average) compared to color. We can test this a couple of ways. 
```{r}
##T-test (can be paired or not)
t.test(color_bw_movies$imdb_score~color_bw_movies$color)

## Wilcoxon T-test (also known as a Mann-Whitney, can be paired or not)
wilcox.test(color_bw_movies$imdb_score~color_bw_movies$color)
```

Let's join the movie data and do some more data exploration
```{r}
movies_joined=full_join(movies_imdb,movies_rottentom,by=c("movie_title"="title"))
```

Let's make a  IMDB, rotten tomatoes, and metacritic scores from the same distribution?

WRITE THE CODE YOURSELVES (density plot? histogram? other ways?)

THEY AREN'T ON THE SAME SCALE, LET'S FIX THAT
use the scale() function

Another way to compare distributions: Quantile-Quantile plots
```{r}
qqplot(movies_joined$imdb_score,movies_joined$rotten_tomatoes)

qqplot(movies_joined$rotten_tomatoes,movies_joined$metacritic)
```

Q-Q plots from the same distribution fall on the x=y line (slope=1, intercept=0) 
```{r}
b=rnorm(1000)
a=rnorm(1000)
qqplot(a,b)
abline(0,1) ## add a line to the plot with an intercept of 0 and a slope of 1
```


Do action movies have a higher IMDB score? 
```{r}
movies_imdb = movies_imdb %>% 
  mutate(action=str_detect(genres,"Action"))
ggplot(movies_imdb,aes(x=imdb_score,fill=action))+geom_density(alpha=0.5)+scale_fill_brewer(palette = "Set1")
``` 

Correlations among variables
```{r}
movies_joined %>% 
  select(imdb_score,rotten_tomatoes,metacritic,budget) %>%
  cor()
``` 

Correlations among variables
```{r}
movies_joined %>% 
  select(imdb_score,rotten_tomatoes,metacritic,budget) %>%
  na.omit() %>%
  cor()
``` 

```{r}
library(corrplot) # or corrgram, up to you
movies_joined %>% 
    select(imdb_score,rotten_tomatoes,metacritic,budget) %>%
    na.omit() %>%
    cor() %>% 
    corrplot(method = "ellipse")
```


Spotting outliers
```{r}
movies_imdb %>% 
  ggplot(aes(x=budget))+
  geom_histogram(bins=20)

movies_imdb %>% 
  ggplot(aes(x=budget))+
  geom_histogram(bins=100)

```


bringing them in closer (log scale)
```{r}
movies_imdb %>% 
  ggplot(aes(x=budget))+
  geom_histogram(bins=100)+
  scale_x_log10()
```


plotting lots of distributions
```{r}
library(ggridges)

movies_imdb %>% 
  ggplot(aes(x=budget,y=color))+
  geom_density_ridges(colour="white",fill="steelblue4")+
  theme_classic(base_size=10) +
  scale_x_log10()
```


it can scale to many distributions
```{r}

movies_imdb %>% 
  ggplot(aes(x=factor(title_year),y=budget))+
  geom_violin(fill="steelblue4")+
  theme_classic(base_size=10) +
  scale_y_log10()

library(ggridges)

movies_imdb %>% 
  ggplot(aes(x=budget,y=factor(title_year)))+
  geom_density_ridges(colour="white",fill="steelblue4")+
  theme_classic(base_size=10) +
  scale_x_log10()

```