```{r}
library(tidyverse)
theme_set(theme_classic(base_size = 20)) 
```

#PCA

install the FactoMineR package, and then load the 'wine' data, and take a subsample (so we don't have too many points)
```{r}
library(FactoMineR)

wine=read_csv("wine/wine_data.csv")

set.seed(10)
wine_sample=wine[sample(1:nrow(wine),500),]

```

This dataset contains 12 colummns of data describing features about wine. Can we describe the  properties of these wines in a smaller number of dimensions?

First limit the data to columns 1-12 and store it as a new object (ignore the wine color for now). We are only working with the numeric columns
```{r}


```

Run a parallel analysis (using fa.parallel() from the psych package) on the remaining wine variables.  How many components would be reasonable to extract?
```{r}

```


Perform a principle components analysis using `prcomp` on these variables and save the results in a new object.  Set the arguments center and scale to be T when running the PCA.
```{r}

```

Print a summary of the object you saved the pca in.  Look at the variance explained by each component.
```{r}

```

Do this in a different (read: easier) way
load in `factoextra` and plot a screeplot of your PCA using `fviz_eig()`
```{r}
library(factoextra)

```

Look at the variable loadings on the first two PCs (the loadings are stored in $rotation).  Look for the highest and lowest loadings and think about what this component might reflect about a wine.  
```{r}

```

look at the variable loadings using the `fviz_pca_var()` function
```{r}

```


Variables loading together on a compoment covary together. To look at this, identify a few variables with the strongest positive loadings on PC1 and look at their correlation with one another.  
```{r}

```

Now look at the correlation (in the raw data) between the variables with the strongest positive loading and the strongest negative loading on PC1
```{r}

```


Plot the scores stored in `.$x` and color the points by the wine color 
```{r}
Can you p

```
If you've made it this far, now try and do the PCA, but without scaling your variables (`scale=F`)
What things are different? What are similar? 