```{r}
library(tidyverse)

theme_set(theme_classic(base_size = 20)) 
```


using lapply to make a list of files (save as object)
```{r}
files=lapply(list.files(path = "movies",full.names = F, pattern = ".txt"),function(file_name){
  read_delim(paste0("movies/",file_name),delim = ",")
})
names(files)=str_replace(string=list.files(path = "movies",full.names = F, pattern = ".txt"),pattern = ".txt",replacement = "")
names(files)
movies_joined=full_join(files[["movies_imdb"]],files[["movies_rottentom"]],by=c("movie_title"="title"))

```

Pirate plot. Arrrrr, from the `ggpirate` package
```{r}
#devtools::install_github("mikabr/ggpirate")
library(ggpirate)

iris %>% 
  ggplot(aes(x=Species,y=Sepal.Width))+
  geom_pirate(aes(colour = Species, fill = Species))


iris %>% ggplot(aes(x=Species,y=Sepal.Width))+
  geom_pirate(aes(colour = Species, fill = Species),bars=F)
```

Antother correlation plot, from the `psych` package
```{r}
library(psych)

movies_joined %>% 
    select(imdb_score,rotten_tomatoes,metacritic,budget) %>%
    pairs.panels()
```


Sampling from a distribution: Binomial
```{r}
cookie_jar=rep.int(x=c("chocolate chip", "oatmeal raisin"),times=c(1000,1000))

sample(x=cookie_jar,size = 10)

library(infer)
tibble(cchip=cookie_jar)  %>% 
  rep_sample_n(size = 10)
```


## sampling >1 time using the "rep_sample_n" function. Here we are sampling 100x
```{r}
tibble(cchip=cookie_jar)  %>% 
  rep_sample_n(size = 10,reps = 100)
```

Now plot
```{r}
tibble(cchip=cookie_jar)  %>% 
  rep_sample_n(size = 10,reps = 100) %>% 
  group_by(replicate) %>% 
  summarize(count=sum(cchip=="chocolate chip")) %>%
  ggplot(aes(x=count))+
  geom_histogram()
```

Can you show me what this looks like if we sample 100 cookies 1000 times?

Now plot with proportions
```{r}
tibble(cchip=cookie_jar)  %>% 
  rep_sample_n(size = 100,reps = 1000) %>% 
  group_by(replicate) %>% 
  summarize(prop=sum(cchip=="chocolate chip")/n()) %>%
  ggplot(aes(x=prop))+
  geom_histogram()
```

Let's take one subsample from our larger population
```{r}
set.seed(7)
cookie_sample=tibble(cchip=cookie_jar)  %>% 
  rep_sample_n(size = 100,reps = 1) 

cookie_sample %>% count(cchip)
```

Let's bootstrap this subsample 1000x to estimate a confidence interval on our point estimate using some other functions in the `infer` package
```{r}
boot_cookie=cookie_sample %>% 
  specify(response = cchip,success = "chocolate chip") %>% 
  generate(reps = 1000,type = "bootstrap") 

boot_cookie %>% 
  calculate(stat = "prop") %>%
  ggplot(aes(x=stat))+
  geom_histogram()
```

calculate a 95% CI
```{r}
boot_cookie %>% 
  calculate(stat = "prop") %>%
  get_ci(level = 0.95, type = "percentile")

## what about the se of the point estimate?
p_cchip=sum(cookie_sample$cchip=="chocolate chip")/length(cookie_sample$cchip)
1.96*sqrt((p_cchip*(1-p_cchip))/length(cookie_sample$cchip))
```

What happens if the true proportion of cookies is closer to 1 or 0? 
```{r}
set.seed(100)
cookie_jar=rep.int(x=c("chocolate chip", "oatmeal raisin"),times=c(950,50))
cookie_sample=tibble(cchip=cookie_jar)  %>% 
  rep_sample_n(size = 100,reps = 1) 

boot_cookie=cookie_sample %>% 
  specify(response = cchip,success = "chocolate chip") %>% 
  generate(reps = 1000) 

boot_cookie %>% 
  calculate(stat = "prop") %>%
  get_ci(level = 0.95, type = "percentile")

## what about the se of the point estimate?
p_cchip=sum(cookie_sample$cchip=="chocolate chip")/length(cookie_sample$cchip)
p_cchip
1.96*sqrt((p_cchip*(1-p_cchip))/length(cookie_sample$cchip))
```

What about when we have a smaller subsample?
```{r}
set.seed(20)
cookie_jar=rep.int(x=c("chocolate chip", "oatmeal raisin"),times=c(950,50))
cookie_sample=tibble(cchip=cookie_jar)  %>% 
  rep_sample_n(size = 25,reps = 1) 

boot_cookie=cookie_sample %>% 
  specify(response = cchip,success = "chocolate chip") %>% 
  generate(reps = 1000) 

boot_cookie %>% 
  calculate(stat = "prop") %>%
  get_ci(level = 0.95, type = "percentile")

## what about the se of the point estimate?
p_cchip=sum(cookie_sample$cchip=="chocolate chip")/length(cookie_sample$cchip)
p_cchip
1.96*sqrt((p_cchip*(1-p_cchip))/length(cookie_sample$cchip))
```

Do movies made before  movies have a higher IMDB rating than romance movies? Can we use a permutation test?

```{r}
movies_action=files[["movies_imdb"]] %>% 
  mutate(action=str_detect(genres,"Action")) %>%
  select(action,title_year,imdb_score)

movies_action%>%
  ggplot(aes(x=action,y=imdb_score))+
  geom_pirate(aes(colour = action),bars = F)+
  scale_color_brewer(palette = "Dark2")
```


```{r}
set.seed(10) ##this allows us to work with the same random sample
movies_action_sample <- movies_action %>% 
  group_by(action) %>%
  sample_n(25) %>% 
  ungroup()

movies_action_sample%>%
  ggplot(aes(x=action,y=imdb_score))+
  geom_pirate(aes(colour = action),bars = F)+
  scale_color_brewer(palette = "Dark2")
```

hypothesis testing with `infer` and using permutations
```{r}
movies_action_sample %>% 
  specify(formula = imdb_score ~ action) %>% 
  calculate(stat = "diff in means", order = c("TRUE", "FALSE"))

movies_action_sample %>% 
  specify(formula = imdb_score ~ action) %>%  ## this shuffles the relationships
  hypothesize(null = "independence") %>% 
  generate(reps = 1,type="permute") %>% 
  calculate(stat = "diff in means", order = c("TRUE", "FALSE"))

movies_action_sample %>% 
  specify(formula = imdb_score ~ action) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 100) %>% 
  calculate(stat = "diff in means", order = c("TRUE", "FALSE"))%>%
  visualise() 
```


compare permutations to bootstrapping
```{r}
permuted=movies_action_sample %>% 
  specify(formula = imdb_score ~ action) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 100,type="permute") %>% 
  calculate(stat = "diff in means", order = c("TRUE", "FALSE"))
permuted$random="permutation"

bootstrapped=movies_action_sample %>% 
  specify(formula = imdb_score ~ action) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 100,type="bootstrap") %>% 
  calculate(stat = "diff in means", order = c("TRUE", "FALSE"))
bootstrapped$random="bootstrap"

combined=rbind(bootstrapped,permuted) #combine them

ggplot(combined,aes(x=stat,fill=random))+
  geom_histogram(position = "dodge")
```

bootstrapping regression coefficients using the `mosiac` package
first we can estimate a regression coefficient
```{r}
set.seed(100) ##this allows us to work with the same random sample
movies_subset=movies_joined %>%
  select(budget,title_year,movie_title,duration,imdb_score,metacritic,rotten_tomatoes) %>%
  sample_n(100)

lm(rotten_tomatoes~imdb_score,movies_subset)

library(broom)  ## to view tidy summaries
tidy(lm(rotten_tomatoes~imdb_score,movies_subset))

```

```{r}
library(mosaic)
lm_boot=do(100) * lm(rotten_tomatoes~imdb_score, 
                     data = sample(movies_subset,replace=T))
confint(lm_boot)
```

what happens if we add in multiple variables?
```{r}
lm_boot=do(100) * lm(rotten_tomatoes~imdb_score + budget, 
                     data = sample(movies_subset,replace=T))
confint(lm_boot)
```








Fisher's exact and Hypergeometric tests (enrichment tests)
``` {r}
MandMs=rep.int(x=c("red", "orange", "yellow", "green", "blue", "brown"),times=c(12, 5, 10, 5, 6, 15))
```

call specific colors from data frame/tibble
```{r}
data.frame(mmcolor=MandMs) %>% 
  ggplot(aes(x=mmcolor,fill=mmcolor))+
  geom_bar() +
  xlab("Color")

data.frame(mmcolor=MandMs) %>% 
  ggplot(aes(x=mmcolor,fill=I(mmcolor)))+
  geom_bar() +
  xlab("Color")
```

A binomial test, but without replacement, so it works well for small sample sizes
```{r}
## let's say we picked out 10 M&Ms. 4 were green and 6 were not green. Did we draw from an unbiased distribution? 
grabbag=tibble(color=rep(x=c( "green", "brown"),times=c(4,6)))

## we'll walk through this together
phyper(4,5,48,5,lower.tail = F)

fisher.test(as.matrix(tibble(green = c(4, 1), notgreen = c(6,48))),alternative = "t")
```


