start your engines (load your packages)
```{r}
library(tidyverse)

theme_set(theme_classic(base_size = 20)) 
```

For loop to read in files (first we list the files)
```{r}
list.files(path = "movies",full.names = F)

list.files(path = "movies",full.names = F, pattern = ".txt")

files=list.files(path = "movies",full.names = F, pattern = ".txt")
for (file_name in files){
  print(file_name)}

for (file_name in files){
  print(paste0("movies/",file_name))}
```

Now that we know what it does, we can read them in. whoops
```{r}
for (file_name in files){
 file=read_delim(paste0("movies/",file_name),delim = ",")
 }
```


can we assign the name of the files? using assign()
```{r}
for (file_name in files){
 assign(file_name,read_delim(paste0("movies/",file_name),delim = ","))
 }
```

can we remove the ".txt" extension from the object name?
```{r}
for (file_name in list.files(path = "movies",full.names = F, pattern = ".txt")){
  file_n=str_replace(file_name,".txt","")
 assign(file_n,read_delim(paste0("movies/",file_name),delim = ","))
}

## clean up the extra objects
rm(file_n)
rm(file_name)
```


using lapply to make a list of files
```{r}
lapply(files,function(file_name){
  read_delim(paste0("movies/",file_name),delim = ",")
})
```

using lapply to make a list of files (save as object)
```{r}
files=lapply(list.files(path = "movies",full.names = F, pattern = ".txt"),function(file_name){
  read_delim(paste0("movies/",file_name),delim = ",")
})

View(files[[1]])
View(files[[2]])

names(files)
names(files)=list.files(path = "movies",full.names = F, pattern = ".txt")
names(files)
names(files)=str_replace(string=list.files(path = "movies",full.names = F, pattern = ".txt"),pattern = ".txt",replacement = "")
names(files)

View(files[["movies_imdb"]])
View(files[["movies_rottentom"]])
```

Sets
```{r}
intersect(movies_imdb$movie_title, movies_rottentom$title) %>% View()
union(movies_imdb$movie_title, movies_rottentom$title) %>% View()
setdiff(movies_imdb$movie_title, movies_rottentom$title) %>% View()
setdiff(movies_rottentom$title, movies_imdb$movie_title) %>% View()

movies_rottentom$title %in% movies_imdb$movie_title %>% View()
```

Let's look at some distributions. 
```{r}
imdb=read_delim("movies/movies_imdb.txt",delim=",")
ggplot(data=movies_imdb,aes(x=imdb_score))+
  geom_histogram()
```

what is a "bin"?
```{r}
ggplot(data=movies_imdb,aes(x=imdb_score))+
  geom_histogram(bins=10)
```

what about "bin width"?
```{r}
ggplot(data=movies_imdb,aes(x=imdb_score))+
  geom_histogram(binwidth = 0.1)
```

What about comparing multiple distributions?
```{r}
ggplot(data=movies_imdb,aes(x=imdb_score,fill=color))+
  geom_histogram(binwidth = 0.1)
```

stacked bar charts are confusing. what else can we do?
```{r}
ggplot(data=movies_imdb,aes(x=imdb_score,fill=color))+
  geom_histogram(binwidth = 0.1, position = "identity")+
  scale_fill_brewer(palette = "Set1")

ggplot(data=movies_imdb,aes(x=imdb_score,color=color))+
  geom_freqpoly(binwidth = 0.1)+scale_color_brewer(palette = "Set1")
```

we have very different sample sizes here, which makes it hard to compare the two distributions. But we could use density plots. 
```{r}
movies_imdb %>% count(color)

ggplot(data=movies_imdb,aes(x=imdb_score,fill=color)) +
  geom_density()+scale_fill_brewer(palette = "Set1")
```

let's get rid of the NAs and make the density plots a bit transparent
```{r}
color_bw_movies=movies_imdb %>% filter(!is.na(color))

ggplot(data=color_bw_movies,aes(x=imdb_score,fill=color))+geom_density(alpha=0.5)+scale_fill_brewer(palette = "Set1")
```

Ohhhh - looks like B/W movies are higher rated (on average) compared to color. We can test this a couple of ways. 
```{r}
##T-test (can be paired or not)
t.test(color_bw_movies$imdb_score~color_bw_movies$color)

## Wilcoxon T-test (also known as a Mann-Whitney, can be paired or not)
wilcox.test(color_bw_movies$imdb_score~color_bw_movies$color)
```

Two variables at a time
First we'll create a new variable for movies being made before or after 2005 (~ half the dataset)
```{r}
movies_imdb %>% filter(!is.na(title_year)) %>% 
  mutate(post2005=title_year>2005) %>% 
  ggplot(aes(x=duration, fill=post2005)) + geom_histogram(alpha=0.5)
movies_imdb %>% filter(!is.na(title_year)) %>%
  mutate(post2005=title_year>2005) %>% 
  ggplot(aes(x=duration, fill=post2005)) + geom_histogram(alpha=0.5, position="identity")
movies_imdb %>% filter(!is.na(title_year)) %>%
  mutate(post2005=title_year>2005) %>% 
  ggplot(aes(x=duration, fill=post2005)) + geom_density(alpha=0.5)
```

What if we have more categories?
we could do boxplots, violinplots, overlapping density plots
note that we're creating a new variable (using mutate) where we cut the title_year variable into 10 equal-sized bins, using "cut()"
```{r}
movies_imdb %>% filter(!is.na(title_year)) %>% 
  mutate(year_bin=cut(title_year,breaks = 10)) %>% 
  ggplot(aes(x=duration, y=year_bin)) + geom_boxplot() +
  ggtitle("boxplot")

movies_imdb %>% filter(!is.na(title_year)) %>% 
  mutate(year_bin=cut(title_year,breaks = 10)) %>% 
  ggplot(aes(x=duration, y=year_bin)) + geom_violin() +
  ggtitle("violin plot")

movies_imdb %>% filter(!is.na(title_year)) %>% 
  mutate(year_bin=cut(title_year,breaks = 10)) %>% 
  ggplot(aes(x=duration, fill=year_bin)) + geom_density(alpha=0.5) +
  ggtitle("density plot")
```

That last one is hard to see, let's look again using ggridges
note that we now have to map the year_bin variable to the y aesthetic, not the fill (you can do both y and fill if you want, but that's redundant)
```{r}
library(ggridges)
movies_imdb %>% filter(!is.na(title_year)) %>% 
  mutate(year_bin=cut(title_year,breaks = 10)) %>% 
  ggplot(aes(x=duration, y=year_bin)) + geom_density_ridges(alpha=0.5)

movies_imdb %>% filter(!is.na(title_year)) %>% 
  mutate(year_bin=cut(title_year,breaks = 10)) %>% 
  ggplot(aes(x=duration, y=year_bin, fill=year_bin)) + geom_density_ridges(alpha=0.5)

movies_imdb %>% filter(!is.na(title_year)) %>% 
    mutate(year_bin=cut(title_year,breaks = 10)) %>% 
    ggplot(aes(x=duration, y=year_bin, fill=year_bin)) +
    geom_density_ridges(alpha=0.5,stat = "binline", bins = 50)

movies_imdb %>% filter(!is.na(title_year)) %>% 
  mutate(year_bin=cut(title_year,breaks = 10)) %>% 
  ggplot(aes(x=duration, y=year_bin, fill=year_bin)) + 
  geom_density_ridges(alpha=0.5, quantile_lines = TRUE, vline_size = 1)

```

How about two continuous variables? 
```{r}
movies_imdb %>% filter(!is.na(title_year)) %>% 
  ggplot(aes(x=duration, y=title_year)) + geom_point()
```

How about many continuous variables?
```{r}
library(GGally)
movies_imdb %>% select(title_year,duration,imdb_score,gross) %>% 
  ggpairs()

movies_imdb %>% select_if(is.numeric) %>% 
  ggpairs()
```






Let's join the movie data and do some more data exploration
```{r}
movies_joined=full_join(movies_imdb,movies_rottentom,by=c("movie_title"="title"))
```

Let's make a  IMDB, rotten tomatoes, and metacritic scores from the same distribution?
```{r}
movies_joined %>% 
    select(movie_title,metacritic,imdb_score,rotten_tomatoes) %>%
    gather(key=source,value = score,c(metacritic,imdb_score,rotten_tomatoes)) %>%
    ggplot(aes(x=score,fill=source))+
    geom_density(alpha=0.5) +
    scale_fill_brewer(palette = "Set1")
```

THEY AREN'T ON THE SAME SCALE, LET'S FIX THAT
use the scale() function
```{r}
movies_joined %>% 
    select(movie_title,metacritic,imdb_score,rotten_tomatoes) %>%
    gather(key=source,value = score,c(metacritic,imdb_score,rotten_tomatoes)) %>%
    group_by(source) %>%
    mutate(score=scale(score)) %>%
    ggplot(aes(x=score,fill=source))+
    geom_density(alpha=0.5) +
    scale_fill_brewer(palette = "Set1")

```
